A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950

### 2022-01-14 paper.Rmd
use lag=0 (no lag)

This is the simplest part, because *B-P* predicts that the baseline distribution (given the log-cube
transformation) is *exponential*, that is, the hazard function is *constant*

```{r postanal, results='asis'}
infdat$parity <- cut(infdat$parity, c(-2, 1.5, 4.5, 19), labels = c("1", "2-4", "5+"))
pneo <- age.window(infdat, c(28 / 365, 1))
##
pneo$enter <- g(pneo$enter * 365)
pneo$exit <- g(pneo$exit * 365)
##
pneo$emintemp <- round(pneo$emintemp)
##
tpneo <- toTpch(Surv(enter, exit, event) ~ sex + socBranch + socStatus + illeg + parity + cold + emintemp + year + week, 
                data = pneo, cuts = g(c(28, 365)))
##fit2 <- phreg(Surv(enter - g(28), exit - g(28), event) ~ sex + socBranch + socStatus + cold + emintemp, 
  ##           data = pneo, shape = 1)
##fit.pch <- pchreg(Surv(enter, exit, event) ~ sex + socBranch + socStatus + cold + emintemp, 
  ##           data = pneo, cuts = g(c(28, 365)))
fit.tpch <- tpchreg(oe(event, exposure) ~ sex + socBranch + socStatus + illeg + parity + I(year - 1923) + cold + emintemp, 
                    data = tpneo, time = age)
dr <- drop1(fit.tpch, test = "Chisq")
ltx(fit.tpch, dr = dr, caption = "Postneonatal mortality", label = "tab:postanal")
```

@@

```{r neoanal, results = 'asis'}
neo <- age.window(infdat, c(0, 28 /365))
neo$exit <- g(neo$exit * 365)
neo$enter <- g(neo$enter * 365)
fit.ph <- coxreg(Surv(enter, exit, event) ~  socStatus + illeg + parity + I(year - 1923) + socBranch  + cold + emintemp,
                 data = neo[neo$sex == "girl", ])
dr <- drop1(fit.ph, test = "Chisq")
ltx(fit.ph, dr = dr, caption = "Neonatal mortality, girls", label = "tab:neonatal")
```

@@

```{r neoanalboys, results = 'asis'}
fit.ph <- coxreg(Surv(enter, exit, event) ~  socStatus + illeg + parity + I(year - 1923) + socBranch  + cold + emintemp,
                 data = neo[neo$sex == "boy", ])
dr <- drop1(fit.ph, test = "Chisq")
ltx(fit.ph, dr = dr, caption = "Neonatal mortality, boys", label = "tab:neonatal")
```

>>>>

This is the simplest part, because *B-P* predicts that the baseline distribution (given the log-cube
transformation) is *exponential*, that is, the hazard function is *constant*.

```{r postanal, results='asis'}
infdat$parity <- cut(infdat$parity, c(-2, 1.5, 4.5, 19), labels = c("1", "2-4", "5+"))
levels(infdat$subreg) <- c("ume", "ske")
infdat$season <- cut(infdat$week, c(-1, 9.5, 22.5, 35.5, 48.5, 52.5), labels = c("winter", "spring", "summer", "fall", "winter"))
infdat$season <- as.factor(infdat$season)
pneo <- age.window(infdat, c(28 / 365, 1))
##
pneo$enter <- g(pneo$enter * 365)
pneo$exit <- g(pneo$exit * 365)
##
pneo$emintemp <- round(pneo$emintemp)
##
tpneo <- toTpch(Surv(enter, exit, event) ~ sex + socBranch + socStatus + illeg + parity + 
                    cold + emintemp + year + season + subreg, 
                data = pneo, cuts = g(c(28, 365)))
##fit2 <- phreg(Surv(enter - g(28), exit - g(28), event) ~ sex + socBranch + socStatus + cold + emintemp, 
  ##           data = pneo, shape = 1)
##fit.pch <- pchreg(Surv(enter, exit, event) ~ sex + socBranch + socStatus + cold + emintemp, 
  ##           data = pneo, cuts = g(c(28, 365)))
fit.tpch <- tpchreg(oe(event, exposure) ~ sex + socBranch + socStatus + illeg + parity + I(year - 1923) + 
                        cold + emintemp + season + subreg, 
                    data = tpneo, time = age)
dr <- drop1(fit.tpch, test = "Chisq")
ltx(fit.tpch, dr = dr, caption = "Postneonatal mortality.", label = "tab:postanal")
```

@@

```{r neoanal, results = 'asis'}
neo <- age.window(infdat, c(0, 28 /365))
neo$exit <- g(neo$exit * 365)
neo$enter <- g(neo$enter * 365)
fit.ph <- coxreg(Surv(enter, exit, event) ~  socStatus + illeg + parity + I(year - 1923) + socBranch  + cold + emintemp + subreg + season,
                 data = neo[neo$sex == "girl", ])
dr <- drop1(fit.ph, test = "Chisq")
ltx(fit.ph, dr = dr, caption = "Neonatal mortality, girls.", label = "tab:neonatal")
```

@@

```{r neoanalboys, results = 'asis'}
fit.ph <- coxreg(Surv(enter, exit, event) ~  socStatus + illeg + parity + I(year - 1923) + socBranch  + cold + emintemp + subreg + season,
                 data = neo[neo$sex == "boy", ])
dr <- drop1(fit.ph, test = "Chisq")
ltx(fit.ph, dr = dr, caption = "Neonatal mortality, boys.", label = "tab:neonatalboys")
```

## Postneonatal mortality, winter and spring

This is the simplest part, because *B-P* predicts that the baseline distribution (given the log-cube
transformation) is *exponential*, that is, the hazard function is *constant*.

```{r postanalw, results='asis'}
##infdat$parity <- cut(infdat$parity, c(-2, 1.5, 4.5, 19), labels = c("1", "2-4", "5+"))
##levels(infdat$subreg) <- c("ume", "ske")
##infdat$season <- cut(infdat$week, c(-1, 9.5, 22.5, 35.5, 48.5, 52.5), labels = c("winter", "spring", "summer", "fall", "winter"))
##infdat$season <- as.factor(infdat$season)
pneo <- age.window(infdat[infdat$season %in% c("winter", "spring"), ], c(28 / 365, 1))
##
pneo$enter <- g(pneo$enter * 365)
pneo$exit <- g(pneo$exit * 365)
##
pneo$emintemp <- round(pneo$emintemp)
##
tpneo <- toTpch(Surv(enter, exit, event) ~ sex + socBranch + socStatus + illeg + parity + 
                    cold + emintemp + year + subreg, 
                data = pneo, cuts = g(c(28, 365)))
##fit2 <- phreg(Surv(enter - g(28), exit - g(28), event) ~ sex + socBranch + socStatus + cold + emintemp, 
  ##           data = pneo, shape = 1)
##fit.pch <- pchreg(Surv(enter, exit, event) ~ sex + socBranch + socStatus + cold + emintemp, 
  ##           data = pneo, cuts = g(c(28, 365)))
fit.tpch <- tpchreg(oe(event, exposure) ~ sex + socBranch + socStatus + illeg + parity + I(year - 1923) + 
                        cold + emintemp + subreg, 
                    data = tpneo, time = age)
dr <- drop1(fit.tpch, test = "Chisq")
ltx(fit.tpch, dr = dr, caption = "Postneonatal mortality, winter and spring.", label = "tab:postanalw")
```

The result in Table \@ref(tab:postanalw) shows that *climate* (`emintemp`) is more important than 
*weather* (`cold`). Moreover, no signs of interaction between weather or climate and the rest of 
covariates (not shown).

## Neonatal mortality, winter and spring

```{r neoanalw, results = 'asis'}
neo <- age.window(infdat[infdat$season %in% c("winter", "spring"), ], c(0, 28 /365))
neo$exit <- g(neo$exit * 365)
neo$enter <- g(neo$enter * 365)
fit.ph <- coxreg(Surv(enter, exit, event) ~  socStatus + illeg + parity + I(year - 1923) + socBranch  + cold + emintemp + subreg,
                 data = neo[neo$sex == "girl", ])
dr <- drop1(fit.ph, test = "Chisq")
ltx(fit.ph, dr = dr, caption = "Neonatal mortality winter and spring, girls.", label = "tab:neonatalw")
```

```{r neoanalboysw, results = 'asis'}
fit.ph <- coxreg(Surv(enter, exit, event) ~  socStatus + illeg + parity + I(year - 1923) + socBranch  + cold + emintemp + subreg,
                 data = neo[neo$sex == "boy", ])
dr <- drop1(fit.ph, test = "Chisq")
ltx(fit.ph, dr = dr, caption = "Neonatal mortality winter and spring, boys.", label = "tab:neonatalboysw")
```
